
\documentclass{beamer}
\usepackage{listings}
\usepackage{graphicx}
\usepackage{wasysym}
\date{\today}
\lstdefinestyle{basic}{
    captionpos=t,%
    basicstyle=\footnotesize\ttfamily,%
    numberstyle=\tiny,%
    numbers=left,%
    stepnumber=1,%
    frame=single,%
    showspaces=false,%
    showstringspaces=false,%
    showtabs=false,%
    %
    keywordstyle=\color{blue},%
    identifierstyle=,%
    commentstyle=\color{gray},%
    stringstyle=\color{magenta}%
}
\title{Your Resume Is Code}
\author{E. Dunham}
\date{2016-11-11}

\begin{document}


\frame{\titlepage}





\begin{frame}
 \frametitle{About Me}



\begin{itemize}
  \item DevOps Engineer at Mozilla Research
\begin{itemize}
  \item First "real job"
\end{itemize}
  \item Open source nerd
  \item Oregonian (currently PDX)
\end{itemize}



\end{frame}
\begin{frame}
 \frametitle{About You}



\begin{itemize}
  \item Student?
  \item Professional in your field of choice?
  \item Other?
\end{itemize}


\end{frame}
\begin{frame}
 \frametitle{Your resume...}



\begin{columns}
\column{0.4\textwidth}

\textbf{IS}
\begin{itemize}
  \item 1-2 pages
  \item First impression
  \item SEO tool
\end{itemize}

\column{0.5\textwidth}

is \textbf{NOT}
\begin{itemize}
  \item Everything you've ever done
  \item Personal details (in US)
  \item Substiute for networking
\end{itemize}
\end{columns}


\end{frame}
\begin{frame}
 \frametitle{A job hunt...}



\begin{itemize}
  \item Network
\begin{itemize}
  \item Where would you want to work?
  \item Where would you not work again?
  \item Why?
\end{itemize}
  \item Share your goals and preferences
\begin{itemize}
  \item Phrase complaints positively
\end{itemize}
  \item Understand referral bonuses
  \item Solicit recruiters
\begin{itemize}
  \item Interview practice makes perfect!
\end{itemize}
  \item Have resume ready
\end{itemize}


\end{frame}
\begin{frame}
 \frametitle{Your Resume's Audiences}



\begin{itemize}
  \item You
  \item Robots
  \item HR
  \item Interviewers
\end{itemize}


\end{frame}
\begin{frame}
 \frametitle{Build a resume like code}



\begin{itemize}
  \item Identify users' needs
  \item Interfaces to other systems
  \item Continuous delivery
\end{itemize}


\end{frame}
\begin{frame}
 \frametitle{Resume users' needs}



\begin{itemize}
  \item You
\begin{itemize}
  \item "Where was I in 2010?"
  \item "What are all the technologies I used there?"
\end{itemize}
  \item Robots
\begin{itemize}
  \item "Does this document contain that buzzword?"
\end{itemize}
  \item HR
\begin{itemize}
  \item "Is this person educated?"
  \item "Do they have the skills the job req says?"
  \item "How can I get in contact with them?"
\end{itemize}
  \item Interviewer
\begin{itemize}
  \item "What has this person built?"
  \item "What kind of questions should I ask them?"
  \item "What technical interests or experiences do we have in common?"
\end{itemize}
\end{itemize}


\end{frame}
\begin{frame}
 \frametitle{Resume's Interfaces}



\begin{itemize}
  \item Use buzzwords from job postings
  \item Reuse content in professional profiles
  \item Accept feedback
\end{itemize}


\end{frame}
\begin{frame}
 \frametitle{Contiuous Delivery}



\begin{itemize}
  \item Resume is living snapshot of career
  \item Resume is finished when career is
  \item Ship it!
\end{itemize}


\end{frame}
\begin{frame}
 \frametitle{Tools you use on code}



\begin{itemize}
  \item Backups!
  \item Version control
  \item Compile/render source into final product that's...
\begin{itemize}
  \item Aesthetically pleasing
  \item Single file
  \item Looks right across platforms
\end{itemize}
  \item Peer review
\end{itemize}


\end{frame}
\begin{frame}
 \frametitle{Version Control}



\begin{itemize}
  \item Backups + searchable history + easily switch between versions
\begin{itemize}
  \item Git, Hg, etc
  \item GitHub, BitBucket, GitLab, whatever
\end{itemize}
  \item Version Control <3 text files
\end{itemize}


\end{frame}
\begin{frame}
 \frametitle{Text File -> PDF}



\begin{itemize}
  \item LibreOffice document?
\begin{itemize}
  \item Hard to version
\end{itemize}
  \item HTML page?
\begin{itemize}
  \item Hard to render nice PDF consistently
\end{itemize}
  \item Write your own typesetting tool
\begin{itemize}
  \item Good experience
  \item Giant rabbit hole
  \item Might never finish your resume
\end{itemize}
  \item Use LaTeX
\begin{itemize}
  \item Useful skill
  \item It has ALL THE FEATURES\footnote{which you find by Googling a lot}
\end{itemize}
\end{itemize}


\end{frame}
\begin{frame}
 \frametitle{LaTeX}



\begin{itemize}
  \item This one time in computer history...
\begin{itemize}
  \item 1969, Donald Knuth write TAOCP
  \item 1976, 2nd edition, phototypesetting
  \item 1978, Knuth publishes TeX
\end{itemize}
  \item Turing-complete language
\begin{itemize}
  \item Expand macros then execute tokens
\end{itemize}
  \item LaTeX == "TeX-based language"
\begin{itemize}
  \item http://www.tug.org/levels.html
\end{itemize}
\end{itemize}


\end{frame}
\begin{frame}
 \frametitle{Learning Most Languages}



\begin{enumerate}
  \item Write some code
  \item Try to run it
  \item Watch it crash
  \item Google a lot
  \item Make it ugly but working
  \item Refine and refactor
\end{enumerate}


\end{frame}
\begin{frame}
 \frametitle{Principles of Resume}



\begin{itemize}
  \item Consistency
  \item Empathy for users' needs
\end{itemize}


\end{frame}
\begin{frame}
 \frametitle{Let's write!}




\end{frame}


\end{document}

